Co-citations in context

### 2019-06-13 cocit_r.tex
sanitized excessive reference to Uzzi

\subsection{Model Misspecification and the Attributes of Disciplinary Context} A source of misspecification in the Uzzi {\em et al.}~model arises from not accounting for disciplinary heterogeneity by treating all references as equiprobable substituents. Under this model, the probability of replacing a reference in one discipline by another reference in a second discipline is identical to the proportion of the articles in WoS that from the second discipline. 
If the Uzzi {\em et al.} model accurately reflects citation practice, the expected proportion of references within papers published in a given discipline $D$ would be equal to the proportion of references in  $D$, and--conversely--the degree to which the proportion deviates from the expected value would reflect the extent of model misspecification.

@@

To study the disciplinary composition of references in our custom datasets, we used the the WoS extended classification consisting of 153 subjects aggregated into the five major research areas of life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities. 
%We measured the frequency with which these five major research areas occur in WoS for the years 1985, 1995, and 2005 and observed the following proportions (some publications qualify for more than one label): (i) life sciences and biomedicine (52.9\%), physical sciences (33.4\%), social sciences (7.2\%), technology (17.9\%), and arts and humanities (1\%). We then matched our applied physics dataset to physical science, and our immunology and metabolism datasets to the life science and biomedicine research areas respectively. We analyzed the composition of the references in our three datasets and observed that 84.2\% of the references in the applied physics dataset belonged to physical sciences, and 93.7\% and 92.1\% of the references in the immunology and metabolism datasets respectively belonged to the life sciences and biomedicine research area. 
The two largest of these research areas are physical sciences and life sciences and biomedicine,  which contribute on average approximately 35.1\% and 62.8\%, respectively, of the references in WoS over the three years of interest.
%While there is some variability in different years, %approximately 35.1\% of the references in WoS are from the physical sciences and approximately 62.8\% of the references are from the life sciences and biomedicine.
While there is some variability over the years, according to the Uzzi  {\em et al.} model, we would expect close to 35.1\% of the references cited by the publications in any large network  to be drawn from the physical sciences and roughly 62.8\% of the references to be drawn from the life sciences and biomedicine.
Yet the empirical data present a very different  story: 79.6\% of the references cited in physical sciences publications are from the physical sciences and 89.9\% of the references cited in life sciences and biomedicine publications are from the life sciences and biomedicine.
In other words, the empirical data shows a strong tendency of publications to cite papers that are in the same major research area rather than in some other research area.  Thus, there is a strong bias towards citations that are {\em inside-the-network} rather than {\em outside-the-network}.
This is inconsistent with the Uzzi {\em et al.} model, but makes sense, given what is well established about citation practice.
%These data suggest at a high level (i) that applying a discipline-indifferent model in Monte Carlo simulations would result in misspecification (ii) strongly that references are likely to belong to the same research area as the publications that cite them rather than being drawn evenly from all five research areas. 
For example, our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, even though that percentage varied among disciplines in the eight specialties they investigated (from approximately 39\% to 89\% for 2006). 
Furthermore, these findings strongly argue that Uzzi {\em et al.}'s model has substantial model misspecification (in deviating substantially from the empirical data), and supports our concern that
definitions of  innovation and conventionality that are based on deviation from expected values under Uzzi's model may not be merited.

%\textcolor{blue}{I don't think the paragraph above makes the point that the distribution that is observed is inconsistent with the Uzzi et al. model. But to make that point, it would be nice to know the proportion of publications in each of the five major subject areas. With that information,  I think this is where we can do a statistical test, and show that these numbers are highly unlikely under the random model.}

@@

We also analyzed the consequences of citation shuffling within a disciplinary set or all of the Web of Science.  References in publications belonging to these three datasets were classified using the WoS Extended Subjects classification of 153 subject areas and summarized as a frequency distribution. A single shuffle of the references in these datasets or all the references in the corresponding WoS year slice was performed, using either the \emph{repcs}  or \emph{umsj} algorithms, after which subject frequencies were computed again. The fold change in subject frequencies of references before and after shuffling was calculated for these groups using all 153 subject categories and are summarized as boxplots in Fig \ref{fig:be}. 
For example, the applied physics dataset contained one reference labeled Genetics and Heredity, but after the shuffle (using the WoS background), there were 1496 references labeled Genetics and Heredity. 
%This proportion was preserved in the intra-network shuffle but increased 1496 times when the WoS background was used as a source of substitution during the Monte Carlo simulation with the \emph{repcs}   algorithm. 
Similarly,  the metabolism dataset  contained one reference labeled Philosophy, but after a single shuffle (again using the WoS background), there were   661 occurrences with this label.
%after a single \emph{repcs}  shuffle. 
The data show convincingly that the disciplinary composition of references in a network is preserved when citation shuffling is constrained to the subnetwork, but is significantly distorted when the WoS superset is used as a source of substitution. A second inference is that the choice between these two algorithms is not relevant (except for running time considerations), and that the background network used for the shuffle is critical.

\begin{table}[ht]
\caption{Model misspecification is reduced by constraining substitutions to disciplinary subnetworks. For the set of journal pairs in common between a disciplinary network and the full WoS dataset, Kullback-Leibler (K-L) divergences
between empirical and simulated journal pair frequencies were computed for the years 1985, 1995, and 2005 for the three disciplinary datasets (applied\_physics, immunology, and metabolism) using either the disciplinary network as background or the WoS superset (all\_wos) to generate the null model (Background). K-L divergence was calculated using the R seewave package \citep{seewave2008} with a base (logarithm) of 2. 
The reduction in K-L divergence between simulated and observed data  shows that model misspecification is reduced by  restricting the reference substitutions to the disciplinary subnetwork (our model), as compared to treating all substitutions within WoS as equiprobable (the Uzzi et al. model).
}
\label{tab:kld}
\centering
\begin{tabular}{| ccccc |} 
  \hline
 Disciplinary Network & Year & Background & K-L Divergence & Ratio \\ 
  \hline
 appl\_physics & 1985 & appl\_physics & 1.21 &  \\ 
  & 1985 & all\_wos & 2.37 & 1.96 \\ 
  & 1995 & appl\_physics & 0.86 &  \\ 
  & 1995 & all\_wos & 2.37 & 2.77 \\ 
  & 2005 & appl\_physics & 0.95 &  \\ 
  & 2005 & all\_wos & 2.35 & 2.47 \\ 
    \hline
  immunology & 1985 & immunology & 0.75 &  \\ 
  & 1985 & all\_wos & 1.68 & 2.24 \\ 
   & 1995 & immunology & 0.78 &  \\ 
   & 1995 & all\_wos & 1.70 & 2.19 \\ 
   & 2005 & immunology & 0.73 &  \\ 
  & 2005 & all\_wos & 1.92 & 2.63 \\ 
    \hline
 metabolism & 1985 & metabolism & 1.11 &  \\ 
   & 1985 & all\_wos & 2.24 & 2.02 \\ 
  & 1995 & metabolism & 1.07 &  \\ 
  & 1995 & all\_wos & 2.33 & 2.17 \\ 
     & 2005 & metabolism & 1.19 &  \\ 
   & 2005 & all\_wos & 2.60 & 2.18 \\ 
   \hline
\end{tabular}
\end{table}

Since our analyses show substantial model misspecification for the Uzzi {\em et al.} model, we then examined a modification to the model that we conjectured would better fit the observed data.  Specifically, 
we conjectured that model misspecification would be reduced by constraining the substitutions to  disciplinary subnetworks,  and tested this by examining the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and predicted citation distributions,
restricted to  the set of journals in a given disciplinary subnetwork.  
The results (Table \ref{tab:kld}) confirm this prediction: simulations under the constrained model (which constrains substitutions to the
disciplinary subnetwork) consistently have a lower K-L divergence compared to simulations under the unconstrained model of Uzzi {\em et al.},  where all references in WoS for a given year are equiprobable.
Furthermore, the K-L divergence for  the unconstrained model is generally twice as large as the K-L divergence for the models where the references are constrained to  the disciplinary subnetworks (a ratio of 1.96 to 2.77,  and  greater than 2.0 for eight out of nine cases).
These results clearly demonstrate that the modification to the Uzzi {\em et al.} model produced by  constraining reference substitutions to the given disciplinary network better models the observed data.

@@

Since the constrained model better fits the observed data, we evaluated the distribution of 
highly cited articles  (i.e., ``hit articles")
 into the four categories (HCHN, LCHN, HCLN, LCLN), for different
thresholds for hit articles.
Figure \ref{fig:Fig2}, Panels (a) and (b), compares hit rates for the four categories among the Immunology, Metabolism, Applied Physics, and WoS datasets for 1995, where the hit rate is defined as the number of hit articles in each category divided by the number of articles in the category. 
The calculation for the hit rates for the WoS dataset (bottom row, Figure \ref{fig:Fig2})
mirrors Uzzi et al.'s results, whereby the largest hit rates were for the HNHC category, despite our methodological improvement of sampling citations in proportion to their frequency. 
However, the trends  for all three  disciplinary subnetworks  are different from those for WoS.
Specifically, the highest hit rates for the 1995 immunology and metabolism datasets are in the LNHC category for the top 1\% of cited articles (and tied between LNHC and HNHC for the top 10\%),
and the highest hit rates for the 1995 applied physics datasets are in the 
HNLC category for both the top 1\% and top 10\% of all cited articles.  
%Across all years and parameter settings, the highest hit rates in the immunology datasets were sometimes in the LNHC category and sometimes in the HNHC category, and the hit rates for the metabolism datasets reflected this same pattern  (Figure \ref{fig:Fig2}). 
 Thus, the category exhibiting the highest hit rate (among highly cited papers) depends on the
specific disciplinary network and to some extent on the threshold for being highly cited.
 %was the case with immunology. 
Furthermore, the categories displaying the greatest hit rate vary 
to some extent % with parameter settings and 
with the year  \textcolor{blue}{(data not shown)}.
%The lack of stable results across parameter settings suggests that parameters must be selected judiciously.

@@

We evaluated the statistical significance of the categorical hit rates using multiple methods, some of which we describe here.  Our first test was based on the null hypotheses that hits were distributed randomly among the four categories with uniform probability in proportion to the number of articles in each category. Using a Chi-Square Goodness of Fit test, rejecting the null hypothesis in favor of the alternate hypothesis supports a non-uniform dispersion of hits: that is, some of the four categories are individually associated with higher than expected, or lower than expected hit rates. The null hypothesis was rejected at a $p<0.001$ in all cases in  Figure \ref{fig:Fig2}, with the exception of the immunology and applied datasets where hit articles are designated as the top 1\% of articles: valid tests were not possible in those instances due to too few expected hits. The null hypothesis was rejected with $p<0.001$ for all valid tests for all parameter settings, all datasets, and all years: hypotheses tests were valid in 73 of 96 instances. We conclude that it is likely that the distribution of hits among categories is not uniform and that instead hit rates vary among the categories in all datasets. 
%(Should we insert a table with all statistical results?)

@@

Our study shows strong evidence that a key assumption of the Uzzi {\em et al.} model that  
substitutions of references  are not impacted by discipline deviates from citation practice, and that
constraining the model to a disciplinary subnetwork reduces the model misspecification.
Furthermore,  using the constrained model instead of the unconstrained model to redefine high (or low) novelty and conventionality produces different trends, depending on the subnetwork.
In particular, while Uzzi {\em et al.} found that highly cited papers were most likely to be
in the HNHC category, this observation does not hold when using the improved model.
Instead, unlike Uzzi {\em et al.}, we find that conventionality flavored with novelty is {\em not} generally a feature of impactful research. 
In particular,   high novelty (as defined by Uzzi {\em et al.}) is
%We conclude that Uzzi et al.'s finding of high hit rates in the HNHC category does not hold generally for disciplinary-based datasets and that 
%novel citation patterns are 
not always indicative of impactful research (as reflected in citation counts), as shown by both Immunology and Metabolism having their highest hit rates for low novelty (Figure \ref{fig:Fig2}).
More generally, these results show that the conclusions of Uzzi {\em et al.} of universal trends in highly cited papers
may be the consequence of using a random model that has a poor fit to the observed data.

Of equal concern, and also contrary to Uzzi et al., we found the category displaying the highest hit rate to be sensitive to the percentage of articles deemed to be hit articles (i.e., the definition of highly cited articles), as well as   and the percentile of articles' z-score distributions that delineated between articles of low novelty and high novelty \textcolor{blue}{(data not shown).}
Faced with this lack of robustness, we are confronted with the necessity of defining defendable parameters, if possible, that determine novelty and conventionality. 
More significantly, we should question whether the simple approach of defining novelty and conventionality, as is done in this article and other research, is sufficient or whether we should expect such variation in the factors associated with impact as we have observed.

%The z-scores that change sign across datasets is either contradictory or an acceptable variation due to the difference in reference sets. We contend that it is the former because the inappropriate substitution of citations with those from disciplines that are implausible and because the observed citation frequencies are ignored.

>>>>

\subsection{Model Misspecification and the Attributes of Disciplinary Context} A source of misspecification in the Uzzi {\em et al.}~model arises from not accounting for disciplinary heterogeneity by treating all references as equiprobable substituents. Under this model, the probability of replacing a reference in one discipline by another reference in a second discipline is identical to the proportion of the articles in WoS that from the second discipline. 
If the Uzzi {\em et al.} model accurately reflects citation practice, the expected proportion of references within papers published in a given discipline $D$ would be equal to the proportion of references in  $D$, and conversely, the degree to which the proportion deviates from the expected value would reflect the extent of model misspecification.

@@

To study the disciplinary composition of references in our custom datasets, we first used the high level WoS classification of five major research areas: life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities. The two largest of these research areas are physical sciences and life sciences and biomedicine, which contribute on average approximately 35.1\% and 62.8\%, respectively, of the references in WoS over the three years of interest. While there is some variability over the years, according to the Uzzi  {\em et al.} model, we would expect close to 35.1\% of the references cited by the publications in any large network  to be drawn from the physical sciences and roughly 62.8\% of the references to be drawn from the life sciences and biomedicine.
Yet the empirical data present a very different  story: 79.6\% of the references cited in physical sciences publications are from the physical sciences and 89.9\% of the references cited in life sciences and biomedicine publications are from the life sciences and biomedicine. In other words, the empirical data shows a strong tendency of publications to cite papers that are in the same major research area rather than in some other research area.  Thus, there is a strong bias towards citations that are {\em intra-network}. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, even though that percentage varied among disciplines in the eight specialties they investigated (from approximately 39\% to 89\% for 2006). Furthermore, these findings argue that a discipline-indifferent random graph model would exhibit misspecification (in deviating substantially from the empirical data), and supports the  concern about definitions of  innovation and conventionality that are based on deviation from expected values.

@@

We also analyzed disciplinary composition at a deeper level of the 153 Subjects in WoS extended classification and examining the consequences of  citation shuffling within a disciplinary set or all of the Web of Science.  References in publications belonging to these three datasets were classified into 153 subject areas and summarized as a frequency distribution. A single shuffle of the references in these datasets or all the references in the corresponding WoS year slice was performed, using either the \emph{repcs}  or \emph{umsj} algorithms, after which subject frequencies were computed again. The fold change in subject frequencies of references before and after shuffling was calculated for these groups using all 153 subject categories and are summarized as boxplots in Fig \ref{fig:be}. For example, the applied physics dataset contained one reference labeled Genetics and Heredity, but after the shuffle (using the WoS background), there were 1496 references labeled Genetics and Heredity. Similarly,  the metabolism dataset  contained one reference labeled Philosophy, but after a single shuffle (again using the WoS background), there were   661 occurrences with this label.
The data show convincingly that the disciplinary composition of references in a network is preserved when citation shuffling is constrained to the subnetwork, but is significantly distorted when the WoS superset is used as a source of substitution. A second inference is that the two algorithms, except for running time considerations, have equivalent effects in this experiment.

\begin{table}[ht]
\caption{Model misspecification is reduced by constraining substitutions to disciplinary subnetworks. For the set of journal pairs in common between a disciplinary network and the full WoS dataset, Kullback-Leibler (K-L) divergences
between empirical and simulated journal pair frequencies were computed for the years 1985, 1995, and 2005 for the three disciplinary datasets (applied\_physics, immunology, and metabolism) using either the disciplinary network as background or the WoS superset (all\_wos) to generate the null model (Background). K-L divergence was calculated using the R seewave package \citep{seewave2008}.  The reduction in K-L divergence between simulated and observed data  shows that model misspecification is reduced by  restricting the reference substitutions to the disciplinary subnetwork (our model), as compared to treating all substitutions within WoS as equiprobable.
}
\label{tab:kld}
\centering
\begin{tabular}{| ccccc |} 
  \hline
 Disciplinary Network & Year & Background & K-L Divergence & Ratio \\ 
  \hline
 appl\_physics & 1985 & appl\_physics & 1.21 &  \\ 
  & 1985 & all\_wos & 2.37 & 1.96 \\ 
  & 1995 & appl\_physics & 0.86 &  \\ 
  & 1995 & all\_wos & 2.37 & 2.77 \\ 
  & 2005 & appl\_physics & 0.95 &  \\ 
  & 2005 & all\_wos & 2.35 & 2.47 \\ 
    \hline
  immunology & 1985 & immunology & 0.75 &  \\ 
  & 1985 & all\_wos & 1.68 & 2.24 \\ 
   & 1995 & immunology & 0.78 &  \\ 
   & 1995 & all\_wos & 1.70 & 2.19 \\ 
   & 2005 & immunology & 0.73 &  \\ 
  & 2005 & all\_wos & 1.92 & 2.63 \\ 
    \hline
 metabolism & 1985 & metabolism & 1.11 &  \\ 
   & 1985 & all\_wos & 2.24 & 2.02 \\ 
  & 1995 & metabolism & 1.07 &  \\ 
  & 1995 & all\_wos & 2.33 & 2.17 \\ 
     & 2005 & metabolism & 1.19 &  \\ 
   & 2005 & all\_wos & 2.60 & 2.18 \\ 
   \hline
\end{tabular}
\end{table}

 We tested the conjecture that model misspecification would be reduced by constraining the substitutions to  disciplinary subnetworks by examining the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and predicted citation distributions, restricted to  the set of journals in a given disciplinary subnetwork.  The results (Table \ref{tab:kld}) confirm this prediction: simulations under the constrained model (which constrains substitutions to the
disciplinary subnetwork) consistently have a lower K-L divergence compared to simulations under the unconstrained model,  where all references in WoS for a given year are equiprobable. Furthermore, the K-L divergence for  the unconstrained model is generally twice as large as the K-L divergence for the models where the references are constrained to  the disciplinary subnetworks (a ratio of 1.96 to 2.77,  and  greater than 2.0 for eight out of nine cases).
These results clearly demonstrate that constraining reference substitutions to the given disciplinary network better models the observed data.

@@

Since the constrained model better fits the observed data, we evaluated the distribution of 
highly cited articles  (i.e., ``hit articles")
 into the four categories (HCHN, LCHN, HCLN, LCLN), for different
thresholds for hit articles.
Figure \ref{fig:Fig2}, Panels (a) and (b), compares hit rates for the four categories among the Immunology, Metabolism, Applied Physics, and WoS datasets for 1995, where the hit rate is defined as the number of hit articles in each category divided by the number of articles in the category. 
The calculation for the hit rates for the WoS dataset (bottom row, Figure \ref{fig:Fig2})
mirrors Uzzi et al.'s results, whereby the largest hit rates were for the HNHC category, despite our methodological changes in sampling citations in proportion to their frequency. 
However, the trends  for all three  disciplinary subnetworks  are different from those for WoS.
Specifically, the highest hit rates for the 1995 immunology and metabolism datasets are in the LNHC category for the top 1\% of cited articles (and tied between LNHC and HNHC for the top 10\%),
and the highest hit rates for the 1995 applied physics datasets are in the 
HNLC category for both the top 1\% and top 10\% of all cited articles.  
%Across all years and parameter settings, the highest hit rates in the immunology datasets were sometimes in the LNHC category and sometimes in the HNHC category, and the hit rates for the metabolism datasets reflected this same pattern  (Figure \ref{fig:Fig2}). 
 Thus, the category exhibiting the highest hit rate (among highly cited papers) depends on the
specific disciplinary network and to some extent on the threshold for being highly cited.
 %was the case with immunology. 
Furthermore, the categories displaying the greatest hit rate vary 
to some extent % with parameter settings and 
with the year  \textcolor{blue}{(data not shown)}.
%The lack of stable results across parameter settings suggests that parameters must be selected judiciously.

@@

We evaluated the statistical significance of the categorical hit rates using multiple methods, some of which we describe here.  Our first test was based on the null hypotheses that hits were distributed randomly among the four categories with uniform probability in proportion to the number of articles in each category. Using a Chi-Square Goodness of Fit test, rejecting the null hypothesis in favor of the alternate hypothesis supports a non-uniform dispersion of hits: that is, some of the four categories are individually associated with higher than expected, or lower than expected hit rates. The null hypothesis was rejected at a $p<0.001$ in all cases in  Figure \ref{fig:Fig2}, with the exception of the immunology and applied datasets where hit articles are designated as the top 1\% of articles: valid tests were not possible in those instances due to too few expected hits. The null hypothesis was rejected with $p<0.001$ for all valid tests for all parameter settings, all datasets, and all years: hypotheses tests were valid in 73 of 96 instances. We conclude that it is likely that the distribution of hits among categories is not uniform and that instead hit rates vary among the categories in all datasets.

@@

Our study shows strong evidence that constraining the model to a disciplinary subnetwork reduces model misspecification. Furthermore,  using the constrained model instead of the unconstrained model to redefine high (or low) novelty and conventionality produces different trends, depending on the subnetwork. In particular, while Uzzi {\em et al.} found that highly cited papers were most likely to be
in the HNHC category, this observation does not consistently hold when using the improved model. Instead, we find that conventionality flavored with novelty is {\em not} generally a feature of impactful research. 
In particular,  high novelty (as defined by Uzzi {\em et al.}) is not always indicative of impactful research (as reflected in citation counts), as shown by both Immunology and Metabolism having their highest hit rates for low novelty (Figure \ref{fig:Fig2}). More generally, these results show that the conclusions of universal trends in highly cited papers may be the consequence of using a random model that has a poor fit to the observed data.

Of relevance, we found the category displaying the highest hit rate to be sensitive to the percentage of articles deemed to be hit articles (i.e., the definition of highly cited articles), and the percentile of articles' z-score distributions that delineated between articles of low novelty and high novelty \textcolor{blue}{(data not shown).} Faced with this lack of robustness, we are confronted with the necessity of defining defendable parameters, if possible, that determine novelty and conventionality.  More significantly, we should question whether the simple approach of defining novelty and conventionality, as is done in this article and other research, is sufficient or whether we should expect such variation in the factors associated with impact as we have observed.
